---
title: "Wastewater-based epidemiology of SARS-CoV-2 in Switzerland"
author: "Julien Riou"
date: "11 August 2023"
output: 
  beamer_presentation:
    colortheme: "spruce"
header-includes: 
  - \setbeamercovered{transparent=30}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# controls
analysis_date = "2023-05-15"
period_dates = c("2022-05-16","2022-09-05","2023-01-02") 
controls = list(analysis_date=analysis_date,
                period_dates=period_dates,
                savepoint=paste0("../savepoints/savepoint_",analysis_date,"/"))
# libs
pacman::p_load(tidyverse,lubridate,ISOweek,INLA,inlabru,sf,cowplot,spdep,scales,flextable)
# source custom functions
fili = dir(path="../R/",pattern="mw_[013456789]",full.names = TRUE)
lapply(X = fili, FUN = function(x) {source(paste0(x), echo=FALSE)})
# load data
ww1 = readRDS(fs::path(controls$savepoint,"ww1.rds"))
shapes = readRDS(fs::path(controls$savepoint,"shapes.rds"))
# aesthetics
theme_set(theme_bw())
theme_update(text = element_text(size=7))
cust_cols = c("dodgerblue","firebrick2","goldenrod2")
```


## Background

Wastewater surveillance of SARS-CoV-2 in Switzerland:

- data available from 7 February 2022
				
- \alert{118 ARAs} (fluctuating)
			
- various sampling frequencies (from weekly to daily)
				
- samples sent to \alert{9 different laboratories}
				
- 20,535 total measurements as of 14 May, 2023


## Data

```{r data_mis1, fig.height=4,fig.width=5.5}
ww1 %>% 
  ggplot() +
    geom_point(aes(x=date,y=ara_kt,colour=kt),size=.3) +
    scale_y_discrete(limits=rev) +
    scale_colour_discrete(guide="none") +
    theme(axis.text.y = element_text(size=3)) +
    labs(x="Date",y="ARA")
```

## Data

```{r data_mis2, fig.height=3.5,fig.width=4.5, fig.cap="Number of viral load measurements by ARA."}
mw_110_map_missing(ww1,shapes)
```

## Data

Large \alert{heterogeneity} across time and space:

```{r data_vl1, fig.height=2.3,fig.width=4,fig.cap="Daily SARS-CoV-2 viral load in wastewater by ARA (removing values below the LOD or LOQ)."}
mw_106_fig_vl_time(ww1)
```

## Objectives

1. Disentangle the \alert{various sources of heterogeneity}

- laboratory, quantification method, systematic temporal or spatial effects, remaining noise... \pause\bigskip
		
2. Extract a clean, ``noise-free'' \alert{temporal signal}

- at the national and/or regional level \pause\bigskip
		
3. Assess the \alert{agreement} with other types of surveillance 

- confirmed cases, hospitalizations, Sentinella, CH-SUR, pooled tests... \pause\bigskip

4. Forecasting/nowcasting 

- historical data, LFO validation \pause\bigskip

5. Future surveillance strategies

- site selection, frequency, rotation

## Methods

\alert{Spatial regression} model using INLA accounting for:

- population covered \pause

- limits of detection (LOD) and of quantification (LOQ) \pause

- \alert{laboratory} and quantification method \pause

- systematic temporal effects (public holidays, weekends) \pause

- \alert{systematic bias} by ARA 

## Methods

Technical aspects:

- gamma likelihood (strictly positive)  \pause

- logarithmic link implying \alert{multiplicative effects} \pause

- iterative model development (model selection tools) \pause

- \alert{random walks} for temporal trends \pause

- (for now) iid for spatial correlation (BYM didn't improve the fit)

<!-- - \alert{Besag-York-Mollié} for spatial correlation (neighbors) -->
			
## Results

Posterior predictive check (\alert{model fit}) is quite good.

```{r res_fit1, fig.height=7,fig.width=12,fig.cap="Model fit."}
ma5.3.2 = readRDS(file=paste0(controls$savepoint,"ma5.3.2.rds"))
ma5.3.3 = readRDS(file=paste0(controls$savepoint,"ma5.3.3.rds"))
ww_all = readRDS(file=paste0(controls$savepoint,"ww_all.rds"))
corr_all_ara = readRDS(file=paste0(controls$savepoint,"corr_all_ara.rds"))
ppp_vl_ara(ww_all,ma5.3.3) + facet_wrap(~ara_name,ncol=10) + labs(title=NULL)
```

## Results

Posterior predictive check (\alert{model fit}) is generally quite good.

```{r res_fit2, fig.height=2.3,fig.width=4,fig.cap="Model fit."}
ppp_vl_ara(ww_all,ma5.3.3,selection=1:6) + facet_wrap(~ara_name,ncol=3) + labs(title=NULL)
```

## Results

Effect of \alert{laboratory} and method (reference is ALTGR_0):

- $\exp(\beta)$ can be interpreted as a \alert{relative viral load}, e.g., the viral load is *on average* 1.43 times higher (0.99 to 2.07) in EAWAG_0 than ALTGR_0
 
```{r res1,  out.width="70%", fig.height=2.1,fig.width=4,fig.cap="Estimated effect of laboratory (laboratory name) and method change (marked by 0 and 1)."}
plot_exp_vl(ma5.3.3,pars="lab|method", ref="ALTGR_0", clean.out = "lab_method") 
```


## Results

Effect of \alert{public holidays} and \alert{weekends}:

- no clear influence
 
```{r res2,  out.width="50%", fig.height=2.1,fig.width=3,fig.cap="Estimated effect of holidays and weekends."}
plot_exp_vl(ma5.3.3,pars="weekend|hol", ref=NULL) + coord_cartesian(ylim=c(.5,2))
```

## Results

Effect of \alert{specific ARAs}:

- some ARAs have consistently higher or lower viral loads

- may be issues with \alert{population} covered (tourism...)
 
```{r res3,  out.width="50%", fig.height=3.5,fig.width=4,fig.cap="Estimated ARA-specific effects."}
mw_130_map_relative_vl(ma5.3.3,corr_all_ara,shapes,forestplot=TRUE, top=10) + 
  labs(title=NULL,y="Relative VL") 
```

## Results

Effect of \alert{specific ARAs}:

- some ARAs have consistently higher or lower viral loads

- may be issues with \alert{population} covered (tourism...)
 
```{r res4,  out.width="80%", fig.height=3.5,fig.width=4,fig.cap="Estimated ARA-specific effects."}
mw_130_map_relative_vl(ma5.3.3,corr_all_ara,shapes,forestplot=FALSE) + 
  labs(title=NULL,y="Relative VL") 
```


## Results

Average temporal trend at the \alert{national} level:

- accounts for all aspects described before

```{r res5,  out.width="80%", fig.height=3,fig.width=5,fig.cap="Estimated average temporal trend at the national level."}
avg_time_trend(ww_all,ma5.3.2) +
  labs(title=NULL,y="Relative VL") 
```


## Results

Average temporal trend at the \alert{regional} level:

- accounts for all aspects described before
 
```{r res6,  out.width="80%", fig.height=3,fig.width=5,fig.cap="Estimated average temporal trend at the regional level."}
avg_time_trend_reg(ww_all,ma5.3.3) +
  labs(title=NULL,y="Relative VL",colour="NUTS2 region") +
  theme(legend.position=c(.4,.18),
        legend.background=element_blank(),
        legend.direction = "horizontal")
```




## Results

\alert{Residual deviations} from the average temporal trend:

- come on top of all aspects described before
 
```{r res7,  out.width="80%", fig.height=3,fig.width=5,fig.cap="Residual deviations from the average temporal trend (top 12 on absolute value)."}
mw_131_map_deviation_from_average(ma5.3.3,corr_all_ara,ww_all,shapes,12)[[1]]  +
  labs(title=NULL,y="Relative VL",colour="NUTS2 region")
```


## Discussion

1. Disentangle the \alert{various sources of heterogeneity}

- important heterogeneity across \alert{laboratories} and \alert{ARAs}

- no clear effect of weekends and public holidays

- possible issue with \alert{population} covered (tourism and/or mistake)

		
## Discussion

2. Extract a clean, ``noise-free'' \alert{temporal signal}

- doable at both the national and regional level

- can look at residuals to identify special situations (Neuchâtel-Jura / Berner Oberland)
		
		
# Discussion

3. Assess the \alert{agreement} with other types of surveillance 

- still ongoing

- idea: joint modelling of reported cases/hospitalisations and viral load

- discussions planned with France (Sorbonne University, Prof. Boëlle) and UK (Imperial, Prof. Blangiardo)





