---
title: "WBE for SARS-CoV-2 in Switzerland: model development (A2) "
author: "Julien Riou"
date: "`r Sys.Date()`"
params:
   controls: controls
output:
   html_document:
     code_folding : hide
     number_sections: false
     highlight: pygments
     theme: cosmo
     link-citations: true
---

```{r results="hide", warnings="false", echo="false"}
source("setup.R")
ww1 = readRDS(fs::path("../",controls$savepoint,"ww1.rds"))
shapes = readRDS(fs::path("../",controls$savepoint,"shapes.rds"))
```


We continue with model A5, now extending to multiple ARAs together. We start with NUTS-2 regions.

# Mittelland

We start with the Mittelland region, that includes 24 ARAs. 


```{r }
# select NUTS2 region
select_NUTS2 = "Mittelland"
ww_reg = ww1 %>%
  # log
  dplyr::mutate(logvl=log(vl)) %>%
  mutate(vl=if_else(vl==0, 1, vl)) %>%
  # select one ARA per NUTS-2
  dplyr::filter(NUTS2_name==select_NUTS2) %>%
  # create indexes for INLA
  dplyr::mutate(day1=day,
                ara1=as.numeric(as.factor(ara_n)),
                ara2=ara1)

# correspondence table
corr_reg = ww_reg %>% 
  group_by(ara_n,ara_id,ara_name,kt,pop,lab,lab_n,ara1,ara2) %>% 
  count() %>% 
  ungroup()

mw_100_desc_table(ww_reg) %>% 
  dplyr::mutate(across(everything(),as.character)) %>% 
  tidyr::gather() %>% 
  flextable::flextable(cwidth=c(4,4)) 

mw_100_desc_table(ww_reg,ara_name) %>% 
  dplyr::select(1:7)  %>% 
  flextable::flextable(cwidth=rep(4,7)) 

```
```{r desc_missing, fig.width=6, fig.height=3.5}
mw_110_map_missing(ww_reg,shapes) + labs(title="Number of measurements")

```
```{r desc_vl, fig.width=8, fig.height=3.5}
ggplot(ww_reg) + geom_line(aes(x=date,y=logvl,colour=ara_name),alpha=.6) + scale_colour_discrete(guide="none") + labs(title="Viral load measurements")
```

 
## Model A5.1: fixed effect
 
We add a geographical structure to model A5, starting by simply adding a fixed effect by ARA, assuming that the time trends of viral load across ARAs are identical. In Mittelland there is no change of methods, and there are 3 laboratories including 2 that only process one ARA each, so we leave these aspects out for now.

```{r ma5.1, fig.width=8, fig.height=8,  R.options = list(width = 1000)}
if(controls$rerun_models) {
  ma5.1 = INLA::inla(vl ~ 1 +
                       f(below_loq,model="iid") +
                       f(below_lod,model="iid") +
                       f(day,model="rw2", scale.model=TRUE, constr=TRUE,
                         hyper=list(prec = list(prior = "pc.prec", param = c(1, 0.01)))) +
                       f(weekend,model="linear",mean.linear=0,prec.linear=.2) +
                       f(hol,model="linear",mean.linear=0,prec.linear=.2) +
                       ara_name,
                     data = ww_reg,
                     family = "gamma",
                     control.compute = list(waic=TRUE,config=TRUE),
                     control.predictor = list(compute=TRUE,link=1))
  saveRDS(ma5.1,file=paste0("../",controls$savepoint,"ma5.1.rds"))
} else {
  ma5.1 = readRDS(file=paste0("../",controls$savepoint,"ma5.1.rds"))
}
summary(ma5.1)
summary_exp_vl(ma5.1,pars="lab|method|hol|weekend")
ppp_vl_ara(ww_reg,ma5.1) 
print("Relative viral load by ARA compared to reference 'Aarwangen (Zala)':")
summary_exp_vl(ma5.1,pars="ara_n",order=TRUE) %>% 
  column_to_rownames(var="Variable")
```

We observe a large heterogeneity across ARAs, with highest relative viral load compared to reference in Lauterbrunnen, Interlaken and Grindelwald and lower relative viral load in Bern, La Chaux-de-Fonds and Emmental (note that viral load already accounts for population size). 
 
## Model A5.2: random intercept
 
We consider using a random intercept by ARA.

```{r ma5.2, fig.width=8, fig.height=8,  R.options = list(width = 1000)}
if(controls$rerun_models) {
  ma5.2 = INLA::inla(vl ~ 1 +
                       f(below_loq,model="iid") +
                       f(below_lod,model="iid") +
                       f(day,model="rw2", scale.model=TRUE, constr=TRUE,
                         hyper=list(prec = list(prior = "pc.prec", param = c(1, 0.01)))) +
                       f(weekend,model="linear",mean.linear=0,prec.linear=.2) +
                       f(hol,model="linear",mean.linear=0,prec.linear=.2) +
                       f(ara1,model="iid"),
                     data = ww_reg,
                     family = "gamma",
                     control.compute = list(waic=TRUE,config=TRUE),
                     control.predictor = list(compute=TRUE,link=1))
  saveRDS(ma5.2,file=paste0("../",controls$savepoint,"ma5.2.rds"))
} else {
  ma5.2 = readRDS(file=paste0("../",controls$savepoint,"ma5.2.rds"))
}
summary(ma5.2)
summary_exp_vl(ma5.2,pars="lab|method|hol|weekend")
ppp_vl_ara(ww_reg,ma5.2) 
print("Relative viral load by ARA compared to average:")
ma5.2$summary.random$ara1 %>% 
  as_tibble() %>% 
  dplyr::transmute(ara1=ID,
                   `exp(beta)`=round(exp(mean),2),
                   `0.025quant`=round(exp(`0.025quant`),2),
                   `0.975quant`=format(round(exp(`0.975quant`),2),scientific=FALSE)) %>% 
  dplyr::left_join(select(corr_reg,ara1,ara_name),by = join_by(ara1)) %>% 
  dplyr::arrange(-`exp(beta)`) %>% 
  select(-ara1) %>% 
  column_to_rownames(var="ara_name")
```

 Results are similar in terms of rankings. Of course the reference is now the between-ARA mean instead of one arbitrary ARA, which is more easily interpreted. As expected ARA-level intercepts are pulled towards the mean.
 
 
## Model A5.3: space-time interaction

We now allow different time trends across ARAs.

```{r ma5.3, fig.width=8, fig.height=8,  R.options = list(width = 1000)}
if(controls$rerun_models) {
  ma5.3 = INLA::inla(vl ~ 1 +
                       f(below_loq,model="iid") +
                       f(below_lod,model="iid") +
                       f(day,model="rw2", scale.model=TRUE, constr=TRUE,
                         hyper=list(prec = list(prior = "pc.prec", param = c(1, 0.01)))) +
                       f(weekend,model="linear",mean.linear=0,prec.linear=.2) +
                       f(hol,model="linear",mean.linear=0,prec.linear=.2) +
                       f(ara1,model="iid") +
                       f(day1,model="rw1", scale.model=TRUE, constr=TRUE,
                         group=ara2, control.group=list(model="iid"),
                         hyper=list(prec = list(prior = "pc.prec", param = c(1, 0.01)))),
                     data = ww_reg,
                     family = "gamma",
                     control.compute = list(waic=TRUE,config=TRUE),
                     control.predictor = list(compute=TRUE,link=1))
  saveRDS(ma5.3,file=paste0("../",controls$savepoint,"ma5.3.rds"))
} else {
  ma5.3 = readRDS(file=paste0("../",controls$savepoint,"ma5.3.rds"))
}
summary(ma5.3)
summary_exp_vl(ma5.3,pars="lab|method|hol|weekend")
ppp_vl_ara(ww_reg,ma5.3)
print("Relative viral load by ARA compared to average:")
ma5.3$summary.random$ara1 %>% 
  as_tibble() %>% 
  dplyr::transmute(ara1=ID,
                   `exp(beta)`=round(exp(mean),2),
                   `0.025quant`=round(exp(`0.025quant`),2),
                   `0.975quant`=format(round(exp(`0.975quant`),2),scientific=FALSE)) %>% 
  dplyr::left_join(select(corr_reg,ara1,ara_name),by = join_by(ara1)) %>% 
  dplyr::arrange(-`exp(beta)`) %>% 
  select(-ara1) %>% 
  column_to_rownames(var="ara_name")
ndays = length(unique(ww_reg$day))
nara = length(unique(ww_reg$ara1))
ma5.3$summary.random$day1 %>% 
  bind_cols(day=rep(0:(ndays-1),nara),
            ara1=rep(1:nara,each=ndays)) %>% 
  left_join(corr_reg,by = join_by(ara1)) %>% 
  ggplot() +
  geom_hline(yintercept=1,linetype=2,alpha=.5) +
  geom_ribbon(aes(x=day,ymin=exp(`0.025quant`),ymax=exp(`0.975quant`),fill=ara_name),alpha=.5) +
  geom_line(aes(x=day,y=exp(mean),colour=ara_name)) +
  facet_wrap(~ara_name) +
  scale_colour_discrete(guide="none") +
  scale_fill_discrete(guide="none") +
  scale_y_continuous(trans="log",breaks = c(.1,1,10)) +
  coord_cartesian(ylim=c(.05,20)) +
  labs(title="Deviations from average time trend by ARA",x="Day",y="Relative viral load by ARA") 
```

This brings a clear improvement in WAIC. We observe similar average between-ARA heterogeneity, with highest relative viral loads in Laupen, Lauterbrunnen, Grindelwald and Neuchâtel (four ARAs where the credible intervals are the highest) and lowest relative viral loads in Delemont, La-Chaux-de-Fonds, Emmental and Bern (four ARAs where the credible intervals are the lowest). While time trends are generally aligned, we observe deviations from the regional average time trend in some ARAs such as Colombier, Delemont and La-Chaux-de-Fonds (comparatively lower viral loads during Summer 2022), and in Lauterbrunnen, Grindelwald and Interlaken (comparatively higher during Summer 2022). 
 
## Model A5.4: spatial structure
 
We now consider the neighboring structure between ARAs.


```{r ma5.4, fig.width=8, fig.height=8,  R.options = list(width = 1000)}
# setup neighboring matrix
shapes_reg = shapes$ara_shp %>% 
  dplyr::filter(ara_id %in% ww_reg$ara_id) %>% 
  left_join(corr_reg,by = join_by(ara_id)) %>% 
  dplyr::arrange(ara1)
sf_use_s2(FALSE)
graph_reg = spdep::poly2nb(shapes_reg)
path_graph = paste0("../",controls$savepoint,"W_reg_",select_NUTS2,".adj")
nb2INLA(path_graph, graph_reg)
if(controls$rerun_models) {
  ma5.4 = INLA::inla(vl ~ 1 +
                       f(below_loq,model="iid") +
                       f(below_lod,model="iid") +
                       f(day,model="rw2", scale.model=TRUE, constr=TRUE,
                         hyper=list(prec = list(prior = "pc.prec", param = c(1, 0.01)))) +
                       f(weekend,model="linear",mean.linear=0,prec.linear=.2) +
                       f(hol,model="linear",mean.linear=0,prec.linear=.2) +
                       f(ara1,model="bym2",
                         graph=path_graph,
                         scale.model = TRUE, constr = TRUE, 
                         hyper = list(theta1 = list("PCprior", c(1, 0.01)),  # Pr(sd<1) = 0.01, unlikely to have rr>3 just based on the spatial confounding
                                      theta2 = list("PCprior", c(0.5, 0.5)))  # Pr(phi<0.5)=0.5, we state that we believe that the unmeasured spatial confounding is driven 50% from the structured and 50% from the unstructured random effect
                       ) +
                       f(day1,model="rw1", scale.model=TRUE, constr=TRUE,
                         group=ara2, control.group=list(model="iid"),
                         hyper=list(prec = list(prior = "pc.prec", param = c(1, 0.01)))),
                     data = ww_reg,
                     family = "gamma",
                     control.compute = list(waic=TRUE,config=TRUE),
                     control.predictor = list(compute=TRUE,link=1))
  saveRDS(ma5.4,file=paste0("../",controls$savepoint,"ma5.4.rds"))
} else {
  ma5.4 = readRDS(file=paste0("../",controls$savepoint,"ma5.4.rds"))
}
summary(ma5.4)
summary_exp_vl(ma5.4,pars="lab|method|hol|weekend")
ppp_vl_ara(ww_reg,ma5.4) 
print("Relative viral load by ARA compared to average:")
ma5.4$summary.random$ara1 %>% 
  as_tibble() %>% 
  dplyr::transmute(ara1=ID,
                   `exp(beta)`=round(exp(mean),2),
                   `0.025quant`=round(exp(`0.025quant`),2),
                   `0.975quant`=format(round(exp(`0.975quant`),2),scientific=FALSE)) %>% 
  dplyr::left_join(select(corr_reg,ara1,ara_name),by = join_by(ara1)) %>% 
  dplyr::arrange(-`exp(beta)`) %>% 
  dplyr::select(-ara1) %>% 
  dplyr::filter(!is.na(ara_name)) %>% 
  column_to_rownames(var="ara_name")
ndays = length(unique(ww_reg$day))
nara = length(unique(ww_reg$ara1))
ma5.4$summary.random$day1 %>% 
  bind_cols(day=rep(0:(ndays-1),nara),
            ara1=rep(1:nara,each=ndays)) %>% 
  left_join(corr_reg,by = join_by(ara1)) %>% 
  ggplot() +
  geom_hline(yintercept=1,linetype=2,alpha=.5) +
  geom_ribbon(aes(x=day,ymin=exp(`0.025quant`),ymax=exp(`0.975quant`),fill=ara_name),alpha=.5) +
  geom_line(aes(x=day,y=exp(mean),colour=ara_name)) +
  facet_wrap(~ara_name) +
  scale_colour_discrete(guide="none") +
  scale_fill_discrete(guide="none") +
  scale_y_continuous(trans="log",breaks = c(.1,1,10)) +
  coord_cartesian(ylim=c(.05,20)) +
  labs(title="Deviations from average time trend by ARA",x="Day",y="Relative viral load by ARA") 
```

 
Results are very similar to model A5.3. We find that the spatial structure explains between 0 and 70% of the spatial heterogeneity.

