---
title: "WBE for SARS-CoV-2 in Switzerland: data description "
author: "Julien Riou"
date: "`r Sys.Date()`"
params:
   controls: controls
output:
   html_document:
     code_folding : hide
     number_sections: false
     highlight: pygments
     theme: cosmo
     link-citations: true
---

```{r results="hide", warnings="false", echo="false"}
source("setup.R")
ww1 = readRDS(fs::path("../",controls$savepoint,"ww1.rds"))
shapes = readRDS(fs::path("../",controls$savepoint,"shapes.rds"))
```

We use measurements of SARS-CoV-2 concentration in wastewater from multiple ARAs in Switzerland in 2022 and 2023. 
Viral concentration (*C*, unit: gene copies [gc] per liter) is transformed into viral load (*V*; unit: gc per day per 100,000) 
using the flow of wastewater on the same day (*F*) and the size of the population covered (*P*):
$$
 V = \frac{C \times F}{P/100,000}m
$$
Table 1 provides a summary of the available data. 
In total 118 ARAs reported data over the full period.
Reporting frequencies and periods varied depending on the ARA, with daily measurements for the full period only available 
in a few cases (Figure 1). ARAs also sent their samples to different laboratories. In some cases, there were also changes 
in the laboratory method used.
SARS-CoV-2 could be detected in the wastewater in most cases, with a few occurrences of no detection (below LOD) and some measurements
below the limit of quantification (LOQ), leading to higher measurement error (Figure 2).
Viral load varied over time, with large heterogeneity across ARAs, although some patterns emerge on visual inspection (Figure 3). 
We also consider covariates at the ARA level. While population is already accounted for in the computation of 
the viral load, covariates included total population covered by the ARA. We also considered the proportion aged below 20 
or above 65, the proportion of non-Swiss non-EU citizen, the average Swiss index of socio-economic position (weighted by hectare),
and the *employment factor* (working places divided by population).


```{r desc_tab}
mw_100_desc_table(ww1) %>% 
  dplyr::mutate(across(everything(),as.character)) %>% 
  tidyr::gather() %>% 
  dplyr::rename(Variable=key,Value=value) %>% 
  flextable::flextable(cwidth=c(4,10)) 
```

**Table 1.** Summary of available data.

```{r desc_tab_nuts2}
mw_100_desc_table(ww1,NUTS2_name) %>% 
  dplyr::select(1:7)  %>% 
  flextable::flextable(cwidth=rep(4,7)) 
```

**Table 2.** Summary of available data by NUTS-2 region.

```{r fig_missing, fig.width=8, fig.height=10}
mw_101_fig_missing(ww1)
```

**Figure 1.** Available measurements over time by ARA (grouped by canton).

```{r map_missing, fig.width=8, fig.height=6}
mw_110_map_missing(ww1,shapes)
```

**Figure 2.** Total measurements by ARA.

```{r fig_detect, fig.width=8, fig.height=10}
mw_103_fig_detect(ww1)
```

**Figure 3.** SARS-CoV-2 detection in wastewater over time by ARA.

```{r fig_vl2, fig.width=8, fig.height=6}
mw_106_fig_vl_time(ww1)
```

**Figure 4.** Daily SARS-CoV-2 viral load in wastewater by ARA (removing values below the LOD or LOQ).

```{r fig_vl, fig.width=8, fig.height=10}
mw_104_fig_vl(ww1)
```

**Figure 5.** Weekly median SARS-CoV-2 viral load in wastewater by ARA (removing values below the LOD or LOQ). Dashed lines show the delimitation in four periods.

```{r map_vl, fig.width=8, fig.height=14}
mw_111_map_vl(ww1,shapes)
```

**Figure 6.** Median SARS-CoV-2 viral load in wastewater by ARA by period.

```{r}
if(FALSE) {
  ggplot(ww1) +
    geom_point(aes(x=pop,y=pop_total)) +
    geom_abline(intercept=0,slope=1)
  
  dd = ww1 %>% 
    group_by(ara_id,ara_name) %>% 
    summarise(pop=max(pop),
              pop_total=max(pop_total)) %>% 
    mutate(rel=(pop_total-pop)/pop,
           abs=pop-pop_total) %>% 
    arrange(-abs(rel))
  ggplot(dd) +
    geom_point(aes(x=ara_name,y=pop),colour="firebrick")+
    geom_point(aes(x=ara_name,y=pop_total),colour="dodgerblue",shape=1) +
    scale_x_discrete(limits=rev) +
    coord_flip() +
  theme(axis.text = element_text(size=5))
}
```

